---
title: "CoViDataIT: CoViD Open Data in Italian Regions"
author: "Leonardo Cefalo"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  \VignetteIndexEntry{CoViDataIT: CoViD Open Data in Italian Regions}
  \VignetteEngine{knitr::rmarkdown}
  \VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(CoViDataIT)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Data source

This dataset relies on the data gathered by Italian regions, validated by the Ministry of Health and eventually shared by the Civil Protection Department. Final open data are available [at this GitHub repository](https://github.com/pcm-dpc/COVID-19).


This dataset is not meant to add any element to the precious information already published by the Civil Protection Department, it is just an `R`-based tool to download automatically the regional level data

## Usage example

At the time being, the usage of the retrieval function `Get_CoViData` is quite simple. The user chooses the time interval across which data are to be downloaded. 

Please notice that no data are available prior to Febryary 24th 2020 neither after January 10th 2025.

If, for instance, the user is interested to get data from January 1st to January 15st 2022, the function call is:

```{r get data}
dd <- Get_CoViData(date_from = "2022-01-01", date_to = "2022-01-15")
str(dd)
```



Notes are quite obnoxious, so here we remove them
```{r}
dd <- dd[, -grep("Notes", names(dd))]
```



```{r data view, echo=FALSE, results='asis'}
knitr::kable(head(dd[-c(1:21),], 10))
```



